---
title: "Homework #5: Customer Retention"
author: "Aditya Arora"
format: pdf
editor: visual
---

## Grading Note

This HW is worth 16 points in total. I've made notes in this document where those points were earned.

## Objectives & setup

-   Homework tasks:
    -   Estimate a model of customer churn from a cell phone carrier -- Data for the homework is contained in `cellphone_churn.csv`, 1 observation (row) per customer -- Customers in the dataset are subscribers to the carrier's unlimited talk plan -- Unlimited plan subscribers have a 1 year contract -- We observe some customer demographics, service usage metrics, and whether the customer cancelled the contract at the end of the term (churn=1) or renewed the service for another year (churn = 0) -- For an active customer, the carrier makes \$40 per month in profit (\$480 per year)
    -   Evaluate the profitability of various retention efforts

The variables in the cellphone_churn.csv are:

| Variable            | Description                                                       |
|--------------------|----------------------------------------------------|
| id                  | Identifier for customer                                           |
| male                | (=1 if customer is male)                                          |
| age                 | Customer age (years)                                              |
| usage               | Average number of minutes of call time per month                  |
| dropped_calls       | Total number of dropped calls (year)                              |
| callcenter_waittime | Total number of minutes spent on hold for customer service (year) |

## Read in the data

To begin, load the customer data into a dataframe named `DF`. Use `head()` and `summary()` to visualize the first few rows and to summarize the variables. **(2 points)**

```{r}
cellphone_churn <- read.csv('cellphone_churn.csv')

head(cellphone_churn)
summary(cellphone_churn)

churn_rate <- (sum(cellphone_churn$churn == 1)/nrow(cellphone_churn))*100
```

**What is the churn rate among the sampled customers? (1 point)**

11.1%

## Question 2

Estimate a model of churn (DV = `churn`) with `male`, `age`, `usage`, `dropped_calls` and `callcenter_waittime` as predictor variables. **(1 point)**

What is the coefficient of `age`, and how do you interpret it? **(1 point)**

```{r}

churn_model <- glm(churn ~ male + age + usage + dropped_calls + callcenter_waittime, data = cellphone_churn, family = binomial)

summary(churn_model)
```

**Answer here**

Coefficient of age = -0.1010797

For every one-unit increase in age, the log odds of churn decrease by approximately 0.1011. Since the coefficient is negative, it suggests that as age increases, the likelihood of churn decreases.

## Question 3

Which of the explanatory variables is the biggest driver of churn (i.e., which variable is responsible for the most variation in predicted churn rates)? **(3 points)**

*Hint: create table of coefficients, delta X, delta prob*

```{r}
# calculate marginal effects

marginal_effects <- coef(churn_model) * c(1, sd(cellphone_churn$male), sd(cellphone_churn$age), sd(cellphone_churn$usage), sd(cellphone_churn$dropped_calls), sd(cellphone_churn$callcenter_waittime))

# delta X = sd(var) if var is continuous, 1 if var is binary
# create table of coefficients,  delta X, delta prob

table_cellphone_churn <- data.frame(
  Coefficients = coef(churn_model),
  Delta_X = c(1, 1, sd(cellphone_churn$age), sd(cellphone_churn$usage), sd(cellphone_churn$dropped_calls), sd(cellphone_churn$callcenter_waittime)),
  Delta_Prob = marginal_effects
)
# note that in model specification, demographic variables are indexed by 4:8

# display table
print(table_cellphone_churn)
```

**Answer here**

Age is the biggest driver of churn since the delta_prob of age is the highest.

## Question 4

Which of the following statements from our analysis of the drivers of churn is TRUE? **Multiple can be TRUE (1 point)**

1.  In the future, acquiring a higher proportion of older, female customers will lead to lower churn rates
2.  In the future, acquiring a higher proportion of younger, female customers will lead to lower churn rates
3.  Reducing call center wait times is the most effective actionable method to reduce churn rates for existing customers.

**Answer here**

Statement 1 and 3 are true.

## Question 5

Calculate the expected CLV of each customer in the sample. For the calculation, assume the following: **(2 points)**

-   annual discount rate = `r` = 0.1
-   CLV horizon = `horizon` = 5 years (0,1,. . . ,4)
-   annual per-customer profit, if retained = `profit` = \$480
-   each customer's retention rate = `RR` = 1 - predicted churn rate
-   retention rates remain constant over time, so use the `CLV_simple()` function we've developed

```{r}
CLV_simple <- function(profit,RR,r,horizon) {
clv = 0
for (t in 0: (horizon-1)) {
clv = clv + profit*(RR/(1+r))^t
}
return(clv)
}

r <- 0.1
profit <- 480
horizon <- 5

cellphone_churn$p.churn <- predict(churn_model, type ="response")
cellphone_churn$retention <- 1 - cellphone_churn$p.churn
cellphone_churn$clv <- 0

for (i in 1:nrow(cellphone_churn)) {
  cellphone_churn$clv[i] = CLV_simple(profit,1-cellphone_churn$p.churn[i],r,horizon)
}
```

What is the average CLV across customers? **(1 point)**

```{r}
mean(cellphone_churn$clv)
```

**Answer here**

\$1696.33

## Question 6

Assume the firm implemented improvements to its network such that `dropped_calls` are reduced by 10%.

What is the average expected increase in CLV as a result of the network improvements? **(2 point)**

*Hint: Use* `predict()` *to generate churn rates assuming* `dropped_calls` *is 10% less than originally observed. Then re-compute CLV values by individual. Finally, compute the difference in average CLV values with network improvements and average CLV values without network improvements.*

```{r}
cellphone_churn2 <- cellphone_churn

cellphone_churn2$dropped_calls <- cellphone_churn$dropped_calls * 0.9 #improved dropped calls

cellphone_churn2$p.churn <- predict(churn_model, type ="response", newdata = cellphone_churn2)

cellphone_churn2$p.retention <- 1 - cellphone_churn2$p.churn

for (i in 1:nrow(cellphone_churn2)) {
  cellphone_churn2$clv[i] <- CLV_simple(profit= 480,RR = cellphone_churn2$p.retention[i],r = 0.1,horizon = 5)
}

Increase.CLV <- mean(cellphone_churn2$clv - cellphone_churn$clv)

print(Increase.CLV)
```

**Answer here**

The average increase in CLV is \$3.03.

## Question 7

Assume the firm implemented improvements to its call center such that `callcenter_waittime` is reduced by 10%. What is the average expected increase in CLV as a result of the call center improvements? **(2 point)**

*Hint: Use* `predict()` *to generate churn rates assuming* `callcenter_waittime` *is 10% less than originally observed. Then re-compute CLV values by individual. Finally, compute the difference in average CLV values with call center improvements and average CLV values without call center improvements.*

```{r}
cellphone_churn3 <- cellphone_churn

cellphone_churn3$callcenter_waittime <- cellphone_churn$callcenter_waittime * 0.9 #improved callcenter waittime

cellphone_churn3$p.churn <- predict(churn_model, type ="response", newdata = cellphone_churn3)

cellphone_churn3$p.retention <- 1 - cellphone_churn3$p.churn

for (i in 1:nrow(cellphone_churn3)) {
  cellphone_churn3$clv[i] <- CLV_simple(profit= 480,RR = cellphone_churn3$p.retention[i],r = 0.1,horizon = 5)
}

Increase.CLV.1 <- mean(cellphone_churn3$clv - cellphone_churn$clv)

print(Increase.CLV.1)
```

**Answer here**

The average increase in CLV is \$8.99.
